Compensated de Casteljau algorithm in \(K\) times the working precision

### 2018-05-24 doc/paper.tex
Hard-coding May 16 in the two documents that used `\today`.
Fixing error in `Split` algorithm in manuscript.
Updating manuscript.
- Adding more specific details when citing a theorem or result
  from another paper, e.g. (\cite{Foo}, Theorem Bar).
- Updating the `tikz_filtration.pdf` image to have one less "row"
  and one less "column" (it was unnecessarily big)

\usepackage{amsthm,amssymb,amsmath}
%% https://tex.stackexchange.com/a/396829/32270
\allowdisplaybreaks
%% H/T: https://tex.stackexchange.com/a/202168/32270
\usepackage{textcomp}
%% H/T: https://tex.stackexchange.com/a/56877/32270
\usepackage{algorithm}
\usepackage{algpseudocode}
%% H/T: https://tex.stackexchange.com/a/244805
\usepackage{booktabs,siunitx}
%% H/T: https://tex.stackexchange.com/a/4217/32270
\usepackage{mathtools}
%% H/T: https://tex.stackexchange.com/a/33869/32270
\makeatletter
\newenvironment{breakablealgorithm}
  {% \begin{breakablealgorithm}
   \begin{center}
     \refstepcounter{algorithm}% New algorithm
     \hrule height.8pt depth0pt \kern2pt% \@fs@pre for \@fs@ruled
     \renewcommand{\caption}[2][\relax]{% Make a new \caption
       {\raggedright\textbf{\ALG@name~\thealgorithm} ##2\par}%
       \ifx\relax##1\relax % #1 is \relax
         \addcontentsline{loa}{algorithm}{\protect\numberline{\thealgorithm}##2}%
       \else % #1 is not \relax
         \addcontentsline{loa}{algorithm}{\protect\numberline{\thealgorithm}##1}%
       \fi
       \kern2pt\hrule\kern2pt
     }
  }{% \end{breakablealgorithm}
     \kern2pt\hrule\relax% \@fs@post for \@fs@ruled
   \end{center}
  }
\makeatother
%% H/T: https://tex.stackexchange.com/a/28334/32270
\usepackage{chngcntr}
\counterwithin{figure}{section}
\counterwithin{table}{section}
\counterwithin{equation}{section}
\counterwithin{algorithm}{section}
%% H/T: https://tex.stackexchange.com/a/202047/32270
%%      https://tex.stackexchange.com/a/32463/32270
\usepackage[labelfont=bf]{caption}
%% H/T: https://tex.stackexchange.com/a/163250/32270
\usepackage{adjustbox}
%% H/T: https://tex.stackexchange.com/a/26348/32270
\renewcommand*{\appendixname}{}

@@

\begin{enumerate}
  \item If \(\delta_i \leq \mach\), \(\rho_i = \pm 1\), then
      \(\prod_{i = 1}^n (1 + \delta_i)^{\rho_i} = 1 + \theta_n\),
  \item \(\left|\theta_n\right| \leq \gamma_n \coloneqq
      n \mach / (1 - n \mach)\),
  \item \((1 + \theta_k)(1 + \theta_j) \leq 1 + \theta_{k + j}\),
  \item \(\gamma_k + \gamma_j + \gamma_k \gamma_j \leq \gamma_{k + j}
    \Longleftrightarrow (1 + \gamma_k)(1 + \gamma_j) \leq 1 + \gamma_{k + j}\),
  \item \((1 + \mach)^j \leq 1 / (1 - j \mach) \Longleftrightarrow
  (1 + \mach)^j - 1 \leq \gamma_j\).
\end{enumerate}

@@

An error-free transformation is a computational method where both
the computed result and the round-off error are returned. It
is considered ``free'' of error if the round-off can be represented
exactly as an element or elements of \(\mathbf{F}\).
The error-free transformations used in this paper are
the \texttt{TwoSum} algorithm by Knuth \cite{Knuth1997} and \texttt{TwoProd}
algorithm by Dekker \cite{Dekker1971}, respectively.

\begin{theorem}[\cite{Ogita2005}]\label{thm:eft}
For \(a, b \in \mathbf{F}\) and \(P, \pi, S, \sigma \in \mathbf{F}\),
\texttt{TwoSum} and \texttt{TwoProd} satisfy
\begin{alignat}{4}
\left[S, \sigma\right] &= \mathtt{TwoSum}(a, b), & \, S &= \fl{a + b},
  S + \sigma &= a + b, \sigma &\leq \mach \left|S\right|,
  & \, \sigma &\leq \mach \left|a + b\right| \\
\left[P, \pi\right] &= \mathtt{TwoProd}(a, b),
  & \, P &= \fl{a \times b}, P + \pi &= a \times b,
  \pi &\leq \mach \left|P\right|,
  & \, \pi &\leq \mach \left|a \times b\right|.
\end{alignat}
The letters \(\sigma\) and \(\pi\) are used to indicate that the
errors came from sum and product, respectively. See
Appendix~\ref{sec:appendix-algo} for implementation details.
\end{theorem}

@@

\begin{theorem}[\cite{Mainar1999}]
If \(p(s) = \sum_{j = 0}^n b_j B_{j, n}(s)\) and \(\mathtt{DeCasteljau}(p, s)\)
is the value computed by the de Casteljau algorithm then
\begin{equation}
\left|p(s) - \mathtt{DeCasteljau}(p, s)\right| \leq \gamma_{2n}
\sum_{j = 0}^n \left|b_j\right| B_{j, n}(s).
\end{equation}
\end{theorem}

@@

\begin{theorem}[\cite{Jiang2010}]
  If no underflow occurs, \(n \geq 2\) and \(s \in \left[0, 1\right]\)
  \begin{equation}
    \frac{\left|p(s) - \mathtt{CompDeCasteljau}(p, s)\right|}{
      \left|p(s)\right|} \leq \mach + 2 \gamma_{3n}^2 \cond{p, s}.
  \end{equation}
\end{theorem}

@@

\begin{figure}
  \includegraphics[width=0.9375\textwidth]{tikz_filtration.pdf}
  \centering
  \captionsetup{width=.75\linewidth}
  \caption{Filtering errors}
  \label{fig:error-filtration}
\end{figure}

@@

\begin{theorem}[\cite{Ogita2005}]\label{thm:sum-k}
A summation can be computed (\texttt{SumK}, Algorithm~\ref{alg:sum-k})
with results that are as accurate as if computed in \(K\) times the
working precision. When computed this way, the result satisfies:
\begin{equation}
\left|\mathtt{SumK}(v, K) - \sum_{j = 1}^n v_j\right| \leq
\left(\mach + 3 \gamma_{n - 1}^2\right) \left|\sum_{j = 1}^n v_j\right| +
\gamma_{2n - 2}^K \sum_{j = 1}^n \left|v_j\right|.
\end{equation}
\end{theorem}

\begin{lemma}[\cite{Jiang2010}]\label{lemma:db-lemma}
The second order error \(\db{2}^{(0)}_0\) satisfies
\begin{equation}
  \left|\db{1}^{(0)}_0 - \cdb{1}^{(0)}_0\right| =
  \left|\db{2}^{(0)}_0\right| \leq 2 \gamma_{3n + 1} \gamma_{3(n - 1)}
  \widetilde{p}(s).
\end{equation}
\end{lemma}

@@

\noindent For IEEE-754 double precision floating point number, \(r = 26\)
so \(2^r + 1\) will be known before \texttt{Split} is called. In all,
\texttt{Split} uses 4 flops.

>>>>

\usepackage{amsthm,amssymb,amsmath}
%% https://tex.stackexchange.com/a/396829/32270
\allowdisplaybreaks
%% H/T: https://tex.stackexchange.com/a/202168/32270
\usepackage{textcomp}
%% H/T: https://tex.stackexchange.com/a/56877/32270
\usepackage{algorithm}
\usepackage{algpseudocode}
%% H/T: https://tex.stackexchange.com/a/244805
\usepackage{booktabs,siunitx}
%% H/T: https://tex.stackexchange.com/a/4217/32270
\usepackage{mathtools}
%% H/T: https://tex.stackexchange.com/a/33869/32270
\makeatletter
\newenvironment{breakablealgorithm}
  {% \begin{breakablealgorithm}
   \begin{center}
     \refstepcounter{algorithm}% New algorithm
     \hrule height.8pt depth0pt \kern2pt% \@fs@pre for \@fs@ruled
     \renewcommand{\caption}[2][\relax]{% Make a new \caption
       {\raggedright\textbf{\ALG@name~\thealgorithm} ##2\par}%
       \ifx\relax##1\relax % #1 is \relax
         \addcontentsline{loa}{algorithm}{\protect\numberline{\thealgorithm}##2}%
       \else % #1 is not \relax
         \addcontentsline{loa}{algorithm}{\protect\numberline{\thealgorithm}##1}%
       \fi
       \kern2pt\hrule\kern2pt
     }
  }{% \end{breakablealgorithm}
     \kern2pt\hrule\relax% \@fs@post for \@fs@ruled
   \end{center}
  }
\makeatother
%% H/T: https://tex.stackexchange.com/a/28334/32270
\usepackage{chngcntr}
\counterwithin{figure}{section}
\counterwithin{table}{section}
\counterwithin{equation}{section}
\counterwithin{algorithm}{section}
%% H/T: https://tex.stackexchange.com/a/202047/32270
%%      https://tex.stackexchange.com/a/32463/32270
\usepackage[labelfont=bf]{caption}
%% H/T: https://tex.stackexchange.com/a/163250/32270
\usepackage{adjustbox}
%% H/T: https://tex.stackexchange.com/a/26348/32270
\renewcommand*{\appendixname}{}
%% H/T: https://tex.stackexchange.com/a/23242/32270
\renewcommand*{\today}{May 16, 2018}

@@

\begin{enumerate}
  \item If \(\delta_i \leq \mach\), \(\rho_i = \pm 1\), then
      \(\prod_{i = 1}^n (1 + \delta_i)^{\rho_i} = 1 + \theta_n\),
  \item \(\left|\theta_n\right| \leq \gamma_n \coloneqq
      n \mach / (1 - n \mach)\),
  \item \((1 + \theta_k)(1 + \theta_j) = 1 + \theta_{k + j}\),
  \item \(\gamma_k + \gamma_j + \gamma_k \gamma_j \leq \gamma_{k + j}
    \Longleftrightarrow (1 + \gamma_k)(1 + \gamma_j) \leq 1 + \gamma_{k + j}\),
  \item \((1 + \mach)^j \leq 1 / (1 - j \mach) \Longleftrightarrow
  (1 + \mach)^j - 1 \leq \gamma_j\).
\end{enumerate}

@@

An error-free transformation is a computational method where both
the computed result and the round-off error are returned. It
is considered ``free'' of error if the round-off can be represented
exactly as an element or elements of \(\mathbf{F}\).
The error-free transformations used in this paper are
the \texttt{TwoSum} algorithm by Knuth (\cite{Knuth1997}) and
\texttt{TwoProd} algorithm by Dekker (\cite{Dekker1971}, Section 5),
respectively.

\begin{theorem}[\cite{Ogita2005}, Theorem 3.4]\label{thm:eft}
For \(a, b \in \mathbf{F}\) and \(P, \pi, S, \sigma \in \mathbf{F}\),
\texttt{TwoSum} and \texttt{TwoProd} satisfy
\begin{alignat}{4}
\left[S, \sigma\right] &= \mathtt{TwoSum}(a, b), & \, S &= \fl{a + b},
  S + \sigma &= a + b, \sigma &\leq \mach \left|S\right|,
  & \, \sigma &\leq \mach \left|a + b\right| \\
\left[P, \pi\right] &= \mathtt{TwoProd}(a, b),
  & \, P &= \fl{a \times b}, P + \pi &= a \times b,
  \pi &\leq \mach \left|P\right|,
  & \, \pi &\leq \mach \left|a \times b\right|.
\end{alignat}
The letters \(\sigma\) and \(\pi\) are used to indicate that the
errors came from sum and product, respectively. See
Appendix~\ref{sec:appendix-algo} for implementation details.
\end{theorem}

@@

\begin{theorem}[\cite{Mainar1999}, Corollary 3.2]
If \(p(s) = \sum_{j = 0}^n b_j B_{j, n}(s)\) and \(\mathtt{DeCasteljau}(p, s)\)
is the value computed by the de Casteljau algorithm then\footnote{In the
original paper the factor on \(\widetilde{p}(s)\) is \(\gamma_{2n}\),
but the authors did not consider round-off when computing
\(1 \ominus s\).}
\begin{equation}
\left|p(s) - \mathtt{DeCasteljau}(p, s)\right| \leq \gamma_{3n}
\sum_{j = 0}^n \left|b_j\right| B_{j, n}(s).
\end{equation}
\end{theorem}

@@

\begin{theorem}[\cite{Jiang2010}, Theorem 5]
  If no underflow occurs, \(n \geq 2\) and \(s \in \left[0, 1\right]\)
  \begin{equation}
    \frac{\left|p(s) - \mathtt{CompDeCasteljau}(p, s)\right|}{
      \left|p(s)\right|} \leq \mach + 2 \gamma_{3n}^2 \cond{p, s}.
  \end{equation}
\end{theorem}

@@

\begin{figure}
  \includegraphics[width=0.875\textwidth]{tikz_filtration.pdf}
  \centering
  \captionsetup{width=.75\linewidth}
  \caption{Filtering errors}
  \label{fig:error-filtration}
\end{figure}

@@

\begin{theorem}[\cite{Ogita2005}, Proposition 4.10]\label{thm:sum-k}
A summation can be computed (\texttt{SumK}, Algorithm~\ref{alg:sum-k})
with results that are as accurate as if computed in \(K\) times the
working precision. When computed this way, the result satisfies:
\begin{equation}
\left|\mathtt{SumK}(v, K) - \sum_{j = 1}^n v_j\right| \leq
\left(\mach + 3 \gamma_{n - 1}^2\right) \left|\sum_{j = 1}^n v_j\right| +
\gamma_{2n - 2}^K \sum_{j = 1}^n \left|v_j\right|.
\end{equation}
\end{theorem}

\begin{lemma}[\cite{Jiang2010}, Theorem 4]\label{lemma:db-lemma}
The second order error \(\db{2}^{(0)}_0\) satisfies\footnote{The authors
  missed one round-off error so used \(\gamma_{3n + 1}\) where
  \(\gamma_{3n + 2}\) would have followed from their arguments.}
\begin{equation}
  \left|\db{1}^{(0)}_0 - \cdb{1}^{(0)}_0\right| =
  \left|\db{2}^{(0)}_0\right| \leq 2 \gamma_{3n + 2} \gamma_{3(n - 1)}
  \widetilde{p}(s).
\end{equation}
\end{lemma}

@@

\noindent For IEEE-754 double precision floating point number, \(r = 27\)
so \(2^r + 1\) will be known before \texttt{Split} is called. In all,
\texttt{Split} uses 4 flops.
